How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985

### 2022-08-23 Draft_June_2022/Full_draft_20220608.rmd
#newplots #newtext #cleanedchunks

---
title: How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article; names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "8 June 2022"
output:
  pdf_document: default
  fig_caption: yes
  html_document:
    df_print: paged
  word_document:
    reference_docx: template.docx
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: The Amazon Rainforest is one of the most important ecosystems in the world, and most of this biome is located in Brazil. Albeit the importance of understanding the relationship between the social construction of the Amazon and its effects on environmental outcomes, we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem in discourses over time, by geographical location, and between, or within, governments. Discursive problem-construction within democracies legitimize ways of thinking and acting towards the environment while affecting decisions to deforest, or not, at the local level. Building on @hirschman1963 and @bacchi2009 conceptualization of chosen problems, we propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in political discourse. Combining hand-coding and supervised machine learning, we classify statements about the Amazon forest or region in 6130 official presidential speeches since 1985. We find that, first, international environmental events appear to drive presidents to speak more about the Amazon. Second, while the Amazon as a problem of economic integration dominated discourses from 1985 to the mid-2000s, environmental conservation and social development constructions briefly surpasses economic integration from 2010 to 2015. Constructing the Amazon as an issue of sovereignty also became more pertinent from 2010 onwards. Finally, using a multinomial regression model, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation, than economic integration or social development, as presidents move away from the Amazon region.
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(readxl)
library(ggplot2)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

#                         location_cat = ifelse(grepl("International", location_cat), 0, 1))

@@

@@

We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem. Historically, the Brazilian government approached the Amazon region, forest and its people as a policy object.Different governmental programs contained an assumptions of what had to be solved in the Amazon, or in other words, it represents the Amazon as a particular problem. These specific problems touch on shared imaginaries of the Amazon as part of the larger socio-cultural history of the country as, for example, in relation to issues of national sovereignty, economic integration, social development, or environmental conservation.

@@

The connection between presidential discourse and the environment has been studied in the case of the United States but remains conceptually underdeveloped. @calderwood2019, for instance, examines 2919 mentions of climate change in American official presidential speeches since 1989. Among various findings, one that stands out is that American presidents frequently side-step the environmental aspects of climate change. He also identifies a shift from economic to security framing of climate issues, side-lining its environmental aspects. @calderwood2020 also tests the effect of geographic location and type of communication regarding climate change. Building prominently on @putnam1988, he hypothesizes that presidents are more likely to mention climate change in foreign locations, and that location influences the specific discursive approach and tone they adopt. The author finds evidence in support of his hypothesis, suggesting American presidential discourse at the top on global warming changes based on location. Elsewhere, @bevitori2015 investigates how the 'environment' has been constructed in American presidential discourse since 1960 using a more automated approach. The author finds that mentions of the environment are typically co-selected with the pronoun 'our', as well as with 'economy', 'clean', and 'preserve'. These studies corroborate the possibilities of variation depending on the audience, but we are yet to tie vocabulary to specific problem-constructions based on wider shared meanings of environment and climate change. While these findings are US specific, they indicate that presidents raise different points about the environment at local, national, or international settings, depending on who they assume their audience is at that specific instance. Nevertheless, as @bacchi2009, we argue that problem-constructions touch on shared meanings about the region that are available to the speaker as part of larger social-cultural history. Mentions to the lexicon as 'environment', 'preserve', or 'climate change' should be tied to the larger meaning of policies in the United States and elsewhere.

@@

# 4 The Amazon in Presidential Speeches

@@

The frequency in which the Amazon appears across all official presidential speeches varies in time. Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches per year. We observe five local maxima: 1989, 1992, 2009, 2015, and 2019. These points, usually, coincide with events that help us explain the rises and falls of the Amazon in presidential discourse in time. For instance, in 1985 the Amazon appeared in about 4% of all the presidential speeches but by 1988 this had increased to around 14%. This is the period when the Brazilian Constitution was written. Indigenous and traditional populations were instrumental in advocating for constitutional environmental rights and the protection of their territories [@hecht1990]. These were eventually enshrined in article 225, which gives all Brazilians a right to a balanced environment, and in article 231, which grants indigenous and traditional populations a right over their territory. However, in 1989, the Amazon appeared in 32% of all speeches. This spike is likely explained by the brutal murder of Chico Mendes and the very end of 1988. An incident that caught unprecedented international attention and Sarney responded to this with a set of policies to address deforestation [@capobianco2021].

@@

perc1 <- pop %>%
  count(amazon_speech, year) %>%
  mutate(perc=n / sum(n)) %>%
  filter(amazon_speech == 1) %>% 
  select(c(perc, amazon_speech, year))
# Simple plot
Amazon_year <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
  geom_line(color = "darkgreen") +
  scale_y_continuous(labels = percent_format()) +
  geom_text(aes(x = 1988.7, y = 0.34, label = "Chico Mendes"), size = 3) +
  geom_text(aes(x = 1992, y = 0.27, label = "Earth Summit"), size = 3) +
  geom_text(aes(x = 2009, y = 0.24, label = "Copenhagen Summit"), size = 3) +
  geom_text(aes(x = 2015, y = 0.21, label = "Paris Summit"), size = 3) +
  geom_vline(xintercept = c(1988.9, 1992, 2009, 2015),
             linetype = "dotted") +
  labs(x = "", y = "", title ="Share of speeches mentioning the Amazon by year",
       subtitle = "Sample composed by 6130 presidential speeches since 1985",
       caption = "",
       color='Y series') +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
Amazon_year
# # comparison with inequality, crime and inflation in speeches
# pop <- pop %>%
#   mutate(desigual = ifelse(grepl("desigual", text, ignore.case = TRUE), 1, 0),
#          crime = ifelse(grepl("violencia|crime|criminalidade|seguranca publica",
#                               text, ignore.case = TRUE), 1, 0),
#          inf = ifelse(grepl("inflacao|inflaçao",
#                              text, ignore.case = TRUE), 1, 0),
#          unemploy = ifelse(grepl("desempreg",
#                              text, ignore.case = TRUE), 1, 0))
# perc2 <- pop %>%
#   count(desigual, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(desigual == 1) %>%
#   select(c(perc, desigual, year))
# perc3 <- pop %>%
#   count(crime, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(crime == 1) %>%
#   select(c(perc, crime, year))
# perc4 <- pop %>%
#   count(inf, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(inf == 1) %>%
#   select(c(perc, inf, year))
# perc5 <- pop %>%
#   count(unemploy, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(unemploy == 1) %>%
#   select(c(perc, unemploy, year))
# colors <- c("Amazon Speech" = "green", "Inequality Speech" = "red",
#             "Criminality Speech" = "yellow", "Inflation Speech" = "blue",
#             "Unemployment Speech" = "purple")
# Amazon_vs_other <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
#   geom_line(aes(color = "Amazon Speech")) +
#   geom_line(data = perc2, aes(x = year, y = perc, color = "Inequality Speech")) +
#   geom_line(data = perc3, aes(x = year, y = perc, color = "Criminality Speech")) +
#   geom_line(data = perc4, aes(x = year, y = perc, color = "Inflation Speech")) +
#   geom_line(data = perc5, aes(x = year, y = perc, color = "Unemployment Speech")) +
#   scale_y_continuous(labels = percent_format()) +
#   scale_color_manual(values = colors) +
#   labs(x = "", y = "", title ="Share of speeches mentioning the Amazon in by year",
#        subtitle = "Sample composed by 6130 presidential speeches since 1985 and compared with mentions of other salient policy objects across the same speeches",
#        caption = "",
#        color = "Legend\n") +
#   theme(panel.background = element_rect("white", "black", .5, "solid"),
#         panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
#         axis.text = element_text(color = "black", size = 10),
#         title = element_text(color = "black", size = 10, face = "bold"),
#         legend.title = element_blank(),
#         plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
#         legend.position = "bottom")
# Amazon_vs_other
# Amazon_vs_other_mean <- tibble::tibble("Policy Object" = c("Amazon", "Inequality",
#                                                            "Criminality", "Inflation",
#                                                            "Unemployment"),
#                                        "Mean share in time" = c(mean(perc1$perc),
#                                                                 mean(perc2$perc),
#                                                                 mean(perc3$perc),
#                                                                 mean(perc4$perc),
#                                                                 mean(perc5$perc)))
# Amazon_vs_other_mean

@@

In 1990 the Amazon appeared in 12% of all speeches and, by 1992, this increased to 25% of all speeches. The 1992 Earth Summit, held in Brazil brought international attention to environmental topics in the Amazon. One of the big announcements, for example, was the consolidation of the first transnational partnership for the Amazon, the G7 Pilot Programme, which brought a high number of financial resources to the region for public policy implementation [@capobianco2021]. Throughout the 1990s and 2000s, during the Cardoso and Lula administrations, the Amazon appeared, on average, on 15% of all speeches. Exceptionally, in 2009 we see another spike in the Amazon appearing in official speeches. This spike coincides with the 2009 Copenhagen Summit as well as the steepest decrease in deforestation rates since data is available. Lula led the delegation to Copenhagen with a self-image of "we do not promise, we deliver" when the stakes about climate change were high [@franchini2019].

@@

We subsequently observe a steady increase from 6% in 2016 to almost 24% in the first year of Bolsonaro's presidency, 2019. At the time, international and national media brought unprecedented attention to Brazilian environmental issues as the record burning of the Amazon, and of the red sky afternoon in São Paulo circulated on social media and international media outlets. This brought issues related to environmental conservation as high priorities in Bolsonaro's governing agenda with efforts to dismantle key pillars of Brazilian environmental governance. At the same time, President Bolsonaro retrieves Brazil's hosting status for COP25. In short, relevant international events drive interest and alter how pressing the "Amazon problem" is perceived to be in society, therefore, cause a momentary shift of interest from top policymakers.

## 4.2 How has the Amazon been constructed as a problem?

@@

We conceptualize four pure problem constructions: sovereignty, economic integration, social development, and conservation. At the level of the speech, though, presidents might mix multiple problem construction in an Amazonian statement. Figure 3, below, illustrates the share of each pure type in time. Pure problem constructions dominate, with an average of 56% of all amazonian statements across time. As above, we observe a strong variation over time for most of these constructions.

```{r Figure 3, fig.cap="Pure types in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mmt1 <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(mx_cat) %>%
  mutate(perc = n / sum(n)) # averages in relation to only mixed-types
mmt2 <- ama_mx %>%
  count(mx_cat)%>%
  mutate(perc = n / sum(n)) # averages in relation to full sample
pure_types_time <- ama_mx %>%
  count(mx_cat, mx_cat2, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(mx_cat2 == "Pure-types") %>% 
  ggplot(aes(x=as.numeric(year), y = perc, color = mx_cat)) +
  geom_smooth(aes(fill = mx_cat), se = FALSE, size = .5) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "", y = "", title ="Share of each pure type by year",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "Curves in the plot were smoothed using loess method.") +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom") +
  ggplot2::annotate(geom = "rect", xmin = hp$beg, xmax = hp$end, ymin = 0, ymax = 0.40,
           fill = c("grey50", "ghostwhite", "grey50", "ghostwhite",
                    "grey50", "ghostwhite", "grey50", "ghostwhite"),
           alpha = 0.1) +
  ggplot2::annotate(geom = "text", x = hp$label, y = 0.01, vjust = 0.3, hjust = 0.3, angle = 25,
           label = hp$presid, alpha = 0.4)
pure_types_time
```

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article; names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "8 June 2022"
output:
  html_document:
    df_print: paged
  fig_caption: yes
  pdf_document: default
  word_document:
    reference_docx: template.docx
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: The Amazon Rainforest is one of the most important ecosystems in the world, and most of this biome is located in Brazil. Albeit the importance of understanding the relationship between the social construction of the Amazon and its effects on environmental outcomes, we lack empirical accounts of how the Brazilian Amazon has been constructed as a problem in discourses over time, by geographical location, and between, or within, governments. Discursive problem-construction within democracies legitimize ways of thinking and acting towards the environment while affecting decisions to deforest, or not, at the local level. Building on @hirschman1963 and @bacchi2009 conceptualization of chosen problems, we propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in political discourse. Combining hand-coding and supervised machine learning, we classify statements about the Amazon forest or region in 6130 official presidential speeches since 1985. We find that, first, international environmental events appear to drive presidents to speak more about the Amazon. Second, while the Amazon as a problem of economic integration dominated discourses from 1985 to the mid-2000s, environmental conservation and social development constructions briefly surpasses economic integration from 2010 to 2015. Constructing the Amazon as an issue of sovereignty also became more pertinent from 2010 onwards. Finally, using a multinomial regression model, we find that presidents are more likely to construct the Amazon as a problem of environmental conservation, than economic integration or social development, as presidents move away from the Amazon region.
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(readxl)
library(ggplot2)
library(gridExtra)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

#                         location_cat = ifelse(grepl("International|Non Identified",
#                                               location_cat), 0, 1))

@@

amazon_in_out <- pop %>% select(amazon_speech, location_cat)
table(amazon_in_out$amazon_speech, amazon_in_out$location_cat)
amazon_in_out <- mutate(amazon_in_out,
                         location_cat = ifelse(grepl("International", location_cat), 0, 1))
amazon_in_out <- table(amazon_in_out$amazon_speech, amazon_in_out$location_cat)
rownames(amazon_in_out) <- c("Not Amazon Speech", "Amazon Speech")
colnames(amazon_in_out) <- c("Outside of Brazil", "Within Brazil")
amazon_in_out
```

@@

We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem. Historically, the Brazilian government approached the Amazon region, forest and its people as a policy object. Different governmental programs contained an assumptions of what had to be solved in the Amazon, or in other words, it represented the Amazon as a particular problem. These specific problems touch on shared imaginaries of the Amazon as part of the larger socio-cultural history of the country as, for example, in relation to issues of national sovereignty, economic integration, social development, or environmental conservation.

@@

The connection between presidential discourse and the environment has been studied in the case of the United States but remains conceptually underdeveloped. @calderwood2019, for instance, examines 2919 mentions of climate change in American official presidential speeches since 1989. Among various findings, one that stands out is that American presidents frequently side-step the environmental aspects of climate change. He also identifies a shift from economic to security framing of climate issues, side-lining its environmental aspects. @calderwood2020 also tests the effect of geographic location and type of communication regarding climate change. Building prominently on @putnam1988, he demonstrates that presidents are more likely to mention climate change in foreign locations, and that location influences the specific discursive approach and tone they adopt. Elsewhere, @bevitori2015 investigates how the 'environment' has been constructed in American presidential discourse since 1960 using a more automated approach. The author finds that mentions of the environment are typically co-selected with the pronoun 'our', as well as with 'economy', 'clean', and 'preserve'. These studies corroborate the possibilities of variation depending on the audience, but we are yet to tie vocabulary to specific problem-constructions based on wider shared meanings of environment and climate change. While these findings are US specific, they indicate that presidents raise different points about the environment at local, national, or international settings, depending on who they assume their audience is at that specific instance. Nevertheless, as @bacchi2009, we argue that problem-constructions touch on shared meanings about the region that are available to the speaker as part of larger social-cultural history. Mentions to the lexicon as 'environment', 'preserve', or 'climate change' should be tied to the larger meaning of policies in the United States and elsewhere.

@@

# 4 How has the Amazon been constructed as a problem?

We break down our analysis in three sub-sections. First, we describe when, where and how often do presidents speak about the Amazon or do not speak about the Amazon. Second, we zoom in speeches about the Amazon and demonstrate how they vary by problem-construction. Finally, we model which specific problem-constructions are more likely to appear in which locations using logistic regressions with multiple controls.

@@

When, where, and how often do presidents speak about the Amazon? Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches in a year. On average, presidents speak about the Amazon in 15% of their speeches. Mentioning the Amazon outside of Brazil is a rarer event than mentioning it inside Brazil.The frequency in which the Amazon appears across presidential speeches (inside, outside, and overall), though, varies in time.

Globally, we observe five local maxima both inside and outside Brazil: 1989, 1992, 2009,2015, and 2019. These points, usually, coincide with events that help us explain the rises and falls of the Amazon in presidential discourse in time. For instance, in 1985 the Amazon appeared in about 4% of all the presidential speeches but by 1988 this had increased to around 14%. This is the period when the Brazilian Constitution was written. Indigenous and traditional populations were instrumental in advocating for constitutional environmental rights and the protection of their territories [@hecht1990]. These were eventually enshrined in article 225, which gives all Brazilians a right to a balanced environment, and in article 231, which grants indigenous and traditional populations a right over their territory. However, in 1989, the Amazon appeared in 32% of all speeches. This spike is likely explained by the brutal murder of Chico Mendes at the very end of 1988. An incident that caught unprecedented international attention and Sarney responded to this with a set of policies to address deforestation [@capobianco2021].

@@

perc1 <- pop %>%
  count(amazon_speech, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>% 
  select(c(perc, amazon_speech, year))

perc_in <- pop %>%
   mutate(location_cat = ifelse(grepl("International|Non Identified", location_cat),
                                "Outside of Brazil", "In Brazil")) %>%
   count(amazon_speech, year, location_cat) %>%
   group_by(year) %>%
   mutate(perc = n / sum(n)) %>%
   filter(amazon_speech == 1) %>%
   select(c(perc, amazon_speech, year, location_cat))

# Plot with All

Amazon_year <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
  geom_line(color = "darkgreen") +
  geom_line(aes(y=mean(perc)), color="black", size=.5)+
  scale_y_continuous(labels = percent_format()) + 
  annotate("text", x = 1988.7, y = 0.34, label = "Chico Mendes", size = 3)+
  annotate("text", x = 1992, y = 0.27, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.24, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.21, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.26, label = "Sao Paulo darkened \n by smoke", size=3) +
  labs(x = "", y = "",
       subtitle = "All speeches inside and outside Brazil (n=6130)",
       caption = "",
       color='Y series') +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "bold"),
        legend.position = "bottom")

# location plots

Amazon_year_in <- perc_in %>% filter(location_cat== "In Brazil") %>% 
  ggplot(., aes(x = year, y = perc))+
  geom_line(color="darkgreen")+
  geom_line(aes(y=mean(perc)), color="black", size=.5)+
  geom_vline(xintercept = c(1988.9, 1992, 2009, 2015, 2019), linetype = "dotted")+
  scale_y_continuous(labels = percent_format())+
  labs(x = "", y = "",
       subtitle = "Speeches in Brazil (n=4416)",
       color='Y series') +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "bold"),
        legend.position = "")

Amazon_year_out <- perc_in %>% filter(location_cat== "Outside of Brazil") %>% 
  ggplot(., aes(x = year, y = perc))+
  geom_line(color="darkgreen")+
  geom_line(aes(y=mean(perc)), color="black", size=.5)+
  geom_vline(xintercept = c(1988.9, 1992, 2009, 2015, 2019), linetype = "dotted")+
  scale_y_continuous(labels = percent_format())+
  labs(x = "", y = "",
       subtitle = "Speeches outside of Brazil (n=1714)",
       caption= "Black lines portray the means.",
       color='Y series') +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "bold"),
        legend.position = "")

grid.arrange(Amazon_year, arrangeGrob(Amazon_year_in, Amazon_year_out, ncol=2), nrow=2, top="Speeches mentioning the Amazon since 1985")

@@

```{r comparison with inequality, crime, and inflation eval=FALSE, include=FALSE}

# # comparison with inequality, crime and inflation in speeches
# pop <- pop %>%
#   mutate(desigual = ifelse(grepl("desigual", text, ignore.case = TRUE), 1, 0),
#          crime = ifelse(grepl("violencia|crime|criminalidade|seguranca publica",
#                               text, ignore.case = TRUE), 1, 0),
#          inf = ifelse(grepl("inflacao|inflaçao",
#                              text, ignore.case = TRUE), 1, 0),
#          unemploy = ifelse(grepl("desempreg",
#                              text, ignore.case = TRUE), 1, 0))
# perc2 <- pop %>%
#   count(desigual, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(desigual == 1) %>%
#   select(c(perc, desigual, year))
# perc3 <- pop %>%
#   count(crime, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(crime == 1) %>%
#   select(c(perc, crime, year))
# perc4 <- pop %>%
#   count(inf, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(inf == 1) %>%
#   select(c(perc, inf, year))
# perc5 <- pop %>%
#   count(unemploy, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(unemploy == 1) %>%
#   select(c(perc, unemploy, year))
# colors <- c("Amazon Speech" = "green", "Inequality Speech" = "red",
#             "Criminality Speech" = "yellow", "Inflation Speech" = "blue",
#             "Unemployment Speech" = "purple")
# Amazon_vs_other <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
#   geom_line(aes(color = "Amazon Speech")) +
#   geom_line(data = perc2, aes(x = year, y = perc, color = "Inequality Speech")) +
#   geom_line(data = perc3, aes(x = year, y = perc, color = "Criminality Speech")) +
#   geom_line(data = perc4, aes(x = year, y = perc, color = "Inflation Speech")) +
#   geom_line(data = perc5, aes(x = year, y = perc, color = "Unemployment Speech")) +
#   scale_y_continuous(labels = percent_format()) +
#   scale_color_manual(values = colors) +
#   labs(x = "", y = "", title ="Share of speeches mentioning the Amazon in by year",
#        subtitle = "Sample composed by 6130 presidential speeches since 1985 and compared with mentions of other salient policy objects across the same speeches",
#        caption = "",
#        color = "Legend\n") +
#   theme(panel.background = element_rect("white", "black", .5, "solid"),
#         panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
#         axis.text = element_text(color = "black", size = 10),
#         title = element_text(color = "black", size = 10, face = "bold"),
#         legend.title = element_blank(),
#         plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
#         legend.position = "bottom")
# Amazon_vs_other
# Amazon_vs_other_mean <- tibble::tibble("Policy Object" = c("Amazon", "Inequality",
#                                                            "Criminality", "Inflation",
#                                                            "Unemployment"),
#                                        "Mean share in time" = c(mean(perc1$perc),
#                                                                 mean(perc2$perc),
#                                                                 mean(perc3$perc),
#                                                                 mean(perc4$perc),
#                                                                 mean(perc5$perc)))
# Amazon_vs_other_mean

# comparison with inequality, crime and inflation in speeches
# pop <- pop %>%
#   mutate(desigual = ifelse(grepl("desigual", text, ignore.case = TRUE), 1, 0),
#          crime = ifelse(grepl("violencia|crime|criminalidade|seguranca publica",
#                               text, ignore.case = TRUE), 1, 0),
#          inf = ifelse(grepl("inflacao|inflaçao",
#                              text, ignore.case = TRUE), 1, 0),
#          unemploy = ifelse(grepl("desempreg",
#                              text, ignore.case = TRUE), 1, 0))
# perc2 <- pop %>%
#   count(desigual, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(desigual == 1) %>%
#   select(c(perc, desigual, year))
# perc3 <- pop %>%
#   count(crime, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(crime == 1) %>%
#   select(c(perc, crime, year))
# perc4 <- pop %>%
#   count(inf, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(inf == 1) %>%
#   select(c(perc, inf, year))
# perc5 <- pop %>%
#   count(unemploy, year) %>%
#   group_by(year) %>%
#   mutate(perc=n / sum(n)) %>%
#   filter(unemploy == 1) %>%
#   select(c(perc, unemploy, year))
# colors <- c("Amazon Speech" = "green", "Inequality Speech" = "red",
#             "Criminality Speech" = "yellow", "Inflation Speech" = "blue",
#             "Unemployment Speech" = "purple")
# Amazon_vs_other <- perc1 %>% ggplot(., aes(x = year, y = perc)) +
#   geom_line(aes(color = "Amazon Speech")) +
#   geom_line(data = perc2, aes(x = year, y = perc, color = "Inequality Speech")) +
#   geom_line(data = perc3, aes(x = year, y = perc, color = "Criminality Speech")) +
#   geom_line(data = perc4, aes(x = year, y = perc, color = "Inflation Speech")) +
#   geom_line(data = perc5, aes(x = year, y = perc, color = "Unemployment Speech")) +
#   scale_y_continuous(labels = percent_format()) +
#   scale_color_manual(values = colors) +
#   labs(x = "", y = "", title ="Share of speeches mentioning the Amazon in by year",
#        subtitle = "Sample composed by 6130 presidential speeches since 1985 and compared with mentions of other salient policy objects across the same speeches",
#        caption = "",
#        color = "Legend\n") +
#   theme(panel.background = element_rect("white", "black", .5, "solid"),
#         panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
#         axis.text = element_text(color = "black", size = 10),
#         title = element_text(color = "black", size = 10, face = "bold"),
#         legend.title = element_blank(),
#         plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
#         legend.position = "bottom")
# Amazon_vs_other
# Amazon_vs_other_mean <- tibble::tibble("Policy Object" = c("Amazon", "Inequality",
#                                                            "Criminality", "Inflation",
#                                                            "Unemployment"),
#                                        "Mean share in time" = c(mean(perc1$perc),
#                                                                 mean(perc2$perc),
#                                                                 mean(perc3$perc),
#                                                                 mean(perc4$perc),
#                                                                 mean(perc5$perc)))
# Amazon_vs_other_mean

In 1990 the Amazon appeared in 12% of all speeches and, by 1992, this increased to 25% of all speeches. The 1992 Earth Summit, held in Brazil brought international attention to environmental topics in the Amazon. One of the big announcements, for example, was the consolidation of the first transnational partnership for the Amazon, the G7 Pilot Programme, which brought a high number of financial resources to the region for public policy implementation [@capobianco2021].Throughout the rest of 1990s and 2000s, during the Cardoso and Lula administrations, the Amazon as a topic did not diverge much from the average. In 2009 we see another spike in the Amazon appearing in official speeches. This spike coincides with the 2009 Copenhagen Summit as well as the steepest decrease in deforestation rates since data is available. Lula led the delegation to Copenhagen with a self-image of "we do not promise, we deliver" when the stakes about climate change were high [@franchini2019].

@@

We subsequently observe a steady increase from 6% in 2016 to almost 24% in the first year of Bolsonaro's presidency, 2019. At the time, international and national media brought unprecedented attention to Brazilian environmental issues as the record burning of the Amazon, and of the red sky afternoon in São Paulo circulated on social media and international media outlets. This brought issues related to environmental conservation as high priorities in Bolsonaro's governing agenda with efforts to dismantle key pillars of Brazilian environmental governance. At the same time, President Bolsonaro retrieves Brazil's hosting status for COP25. In short, relevant events drive interest and alter how pressing the "Amazon problem" is perceived to be in society, therefore, events cause a momentary shift of interest.These events seem to drive the topic up inside and outside Brazil.

## 4.2 Problem-constructions in time

@@

When presidents speak about the Amazon, what are they saying? We conceptualize four pure problem constructions: sovereignty, economic integration, social development, and conservation. At the level of the speech, though, presidents might mix multiple problem construction in an Amazonian statement. Figure 3, below, illustrates the share of each pure type in time. Pure problem constructions dominate, with an average of 56% of all amazonian statements across time. As above, we observe a strong variation over time for most of these constructions.

```{r Figure 3, fig.cap="Pure types in time", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mmt1 <- ama_mx %>%
  filter(mx_cat2 == "Mixed-types") %>%
  count(mx_cat) %>%
  mutate(perc = n / sum(n)) # averages in relation to only mixed-types
mmt2 <- ama_mx %>%
  count(mx_cat)%>%
  mutate(perc = n / sum(n)) # averages in relation to full sample
pure_types_time <- ama_mx %>%
  count(mx_cat, mx_cat2, year) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>% 
  filter(mx_cat2 == "Pure-types") %>% 
  ggplot(aes(x=as.numeric(year), y = perc, color = mx_cat)) +
  geom_smooth(aes(fill = mx_cat), se = FALSE, size = .5) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "", y = "", title ="Share of each pure-type problem-construc by year",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "Curves in the plot were smoothed using loess method.") +
  theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom") +
  ggplot2::annotate(geom = "rect", xmin = hp$beg, xmax = hp$end, ymin = 0, ymax = 0.40,
           fill = c("grey50", "ghostwhite", "grey50", "ghostwhite",
                    "grey50", "ghostwhite", "grey50", "ghostwhite"),
           alpha = 0.1) +
  ggplot2::annotate(geom = "text", x = hp$label, y = 0.01, vjust = 0.3, hjust = 0.3, angle = 25,
           label = hp$presid, alpha = 0.4)
pure_types_time
```
